---
title: How does tree harvest influence benthic invertebrate density?
author:
  - name: Sam Straus
    email: samantha.straus@mcgill.ca
    affiliation: McGill University
    correspondingauthor: true
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
abstract: |
  We looked at the density of Baetis spp. in harvested and non-harvested stream catchments.

keywords: 
  - stream ecology
  - entomology
journal: "Ecological Entomology"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
header-includes:
 \usepackage{float}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r setup, include=FALSE}

library("renv")
library("palmerpenguins")
library("tidyverse")
library(ggplot2)
library(cowplot)

#renv::init()
```

# Introduction

*Baetis* spp, belonging to the insect order Ephemeroptera, are commonly used as stream quality indicators @wallace1986response.

# Methods

Researchers collected data on the number of *Baetis* spp. per square meter in the benthic environment of various catchments from 1995 to 2005. Some of the catchments were harvested, while others were not.

# Results

Figure \ref{fig1} is generated using an R chunk.

```{r, include=FALSE}

dat <- read.csv("../data/TLW_invertebrateDensity.csv")

table(dat$catchment)

## catchments 31, 33, and 34 were the ones that experienced forest harvest, create new column
dat <- dat %>% mutate(harvest = if_else(catchment == "33" | catchment == "34L" | catchment == "34M" | 
                                   catchment == "34U", true = "yes", false = "no"))

table(dat$harvest)

# for the purposes of this mock assignment, let's just stick on with one genus: Baetis (a genus of mayfly, the order of which are stream quality indicators 
head(dat)
baetis <- dat %>% select(catchment, year, month, replicate, Baetis, harvest)

baetis.sum <- baetis %>% group_by(catchment, year, month, harvest) %>% summarize(mean.dens = mean(Baetis))


## need date format to plot as a time series
table(baetis.sum$month)
baetis.sum <- baetis.sum %>% mutate(month2 = if_else(month == "june", "06", if_else(month == "may", "05", 
                      if_else(month == "november", "11", if_else(month =="october", "10",
                      if_else(month == "september", "09", "0"))))))

# unite month and year
baetis.sum <- baetis.sum %>% unite(col = "year.month", year, month2, sep = "-")

## adding a day to get into yyyy-mm-dd format that R likes, the data file does not specify the day, so we're just going to use the first of the month

baetis.sum <- baetis.sum %>% mutate(ymd = as.Date(paste(year.month,"-01",sep="")))

## double check that are actually in a date format
str(baetis.sum$ymd) # nice!

p <- ggplot(baetis.sum, aes(x=ymd, y=mean.dens, color = harvest)) +
  geom_line() + 
  xlab("Year")+
  ylab("Mean density (# per sq. m)")+
  theme_cowplot()
  
p

```

```{r fig1, fig.width = 5, fig.height = 3, fig.align='center', out.width="75%", fig.cap = "\\label{fig1}Baetis density in harvested and non-harvested catchments", echo = FALSE}
p
```

# Discussion

*Baetis* sp. (Ephemeroptera) are used as stream quality indicators for catchments @wallace1986response.

# References {.unnumbered}
